\chapter{CST Macros}
\label{cha:cstmacro}
\def\pil{$\rightarrow$\xspace}
This chapter contains macros for automating CST Microwave Studio. In order to use a macro, go to Macros \pil Make VBA Macro \pil ``MacroName'' \pil Paste the macro code.

\section{Sweep Total Efficiency}
The following macro makes it possible to sweep a parameter (used in Schematic View) and save the System Total Efficiency to a file for each value. In order to use the script, the following must first be done:
\begin{itemize}
    \item Create a farfield monitor in 3D View for each frequency of interest.
    \item In Schematic View, create an AC, Combine results-task. Use this to obtain farfields which are updated when the schematic is updated. This creates ``AC''-versions of the farfields and efficiencies in 3D View.
\end{itemize}
The macro creates a graphical user interface where the user must input
\begin{description}
    \item[Parameter] The parameter to be swept.
    \item[From] The initial parameter value.
    \item[To] The final parameter value.
    \item[Step] The step-size.
    \item[Output dir] The directory where the output files will be stored (should \emph{not} end on \texttt{/}).
    \item[AC] The AC parameter to be stored (e.g.\ AC1).
\end{description}
The macro is shown in Listing~\ref{lst:cstmacro_eff}. The output is one text file for each parameter, located in the given output directory.

\begin{lstlisting}[language={[Visual]Basic}, caption={Sweep Total Efficiency macro for CST.}, label=lst:cstmacro_eff]
' ExportEfficiency

Sub Main ()
    Dim Studio As Object

    Begin Dialog UserDialog 360,161,"Export Efficiency Sweep" ' %GRID:10,7,1,1
        Text 0,0,180,14,"Parameter",.LParameter
        Text 0,21,180,14,"From",.LFrom
        Text 0,42,180,14,"To",.LTo
        Text 0,63,180,14,"Step",.LStep
        Text 0,84,180,14,"Output dir",.LOutputdir
        Text 0,105,180,14,"AC (e.g. AC1, AC2, ...)",.LAc
        TextBox 180,0,180,21,.EParameter
        TextBox 180,21,180,21,.EFrom
        TextBox 180,42,180,21,.ETo
        TextBox 180,63,180,21,.EStep
        TextBox 180,84,180,21,.EOutputDir
        TextBox 180,105,180,21,.EAc
        OKButton 180,140,90,21
        CancelButton 270,140,90,21
    End Dialog
    Dim dlg As UserDialog
    Dialog dlg

    ' Do parameter sweep. Save txt file for each iteration
    Dim ThisVal As Double
    For ThisVal = CDbl(dlg.EFrom) To CDbl(dlg.ETo)+0.01 STEP CDbl(dlg.EStep)
        StoreParameter(dlg.EParameter, ThisVal)
        DS.UpdateResults
        SelectTreeItem "1D Results\Efficiencies\System Tot. Efficiency [" + dlg.EAc + "]"
        ExportPlotData dlg.EOutputDir + "\" + CStr(ThisVal) + ".txt"
    Next
End Sub
\end{lstlisting}
